### 2015-07-12 narrative/data-reduction.tex

Deep application knowledge means one can sometimes achieve drastically superior
reduction in the total data compared with what one might achieve otherwise.
However, even in the absence of such high level knowledge, we must equip the
SSIO layers with generic data reduction techniques and re-organization
techniques. For instance, certain basic information on the semantics of the
data is needed and must be supported by the overall infrastructure. Likewise,
information regarding typical access patterns for data in a given application
should be placed in the SSIO layers. Scientific applications often read and
analyze data in a structured fashion and failure to respect this pattern, or to
exploit it through prefetching, can significantly increase the overall
execution time of the application. Accurate prefetching can only be achieved
with access pattern analysis. This requires an SSIO layer which is capable of
prefetching data in an asynchronous fashion to take the bytes on the file.

@@

\item Precision based re-organization. This is where the most significant bytes
	of the data are all grouped together from each object. This data
	generally will have a higher utility then the data with the least
	significant bytes.  This process has the following three steps which
	need to be efficiently implemented when data gets placed to the storage
	system and from the storage system.  First, the data needs to be
	re-arranged which involves memory operations and needs to be done
	in situ. This operation involves no communication.  Data needs to be
	copied, so the memory requirements are increased. At most this requires
	two copies of an individual dataset in memory, and we will investigate
	techniques to allocate and deallocate this memory if the user will
	specify that the data will be overwritten after it is written to the
	storage system. This is often the case for many of the  quantities
	written from the simulation, but there are many cases where, for
	example, we want to write all of the particles from a Particle In Cell
	(PIC) simulation.  Since the particles will be used later in the
	calculation, we need to duplicate the storage. Our observation with
	working with the XGC1, GTC, and PiconGPU simulations 
	 is that we can temporary increase the
	storage of the particles, and then release them since the temporary
	arrays used in the calculations are often freed when a PIC iteration is
	finished. 
%
\item Frequency-based re-organization.  Another common technique to classify
	the importance of data is to first re-organize the data in frequency
	space and then specify that as the frequency increases the importance
	of data decreases.   Common techniques  used for streaming data,
	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
	code streams which have regions of interest that offer several
	mechanisms to support spatial random access or region of interest
	access at varying degrees of granularity. It is possible to store
	different parts of the same data using different quality.  This allows
	us to place the lowest frequency chunks in the fastest storage, and the
	highest frequency chunks  on either  the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out  these pieces. 
%	This binning of data in frequency space can take as an input
%	(storage  {\color{red}ending )? CHECK-tsr} bandwidth knowledge, storage size limitations, and user
%	intentions of how long the majority of the data will be frequently
%	accessed. 
	 In order to fully take advantage of frequency based
	re-organization we have devised a scheme where the data pre-conditioner
	first sorts the data in bins of 1024 words, then we maintain the
	sort order, and then we used wavelets and spline fits to then
	re-organize the values.  This re-organization allows us to understand
	that the resultant data is smoother now, and can be reduced in each one
	of the bins.  \end{itemize}

>>>>

Deep application knowledge means one can sometimes achieve drastically superior
reduction in the total data compared with what one might achieve otherwise.
However, even in the absence of such high level knowledge, we must equip the
SSIO layers with generic data reduction techniques and re-organization
techniques. For instance, certain basic information on the semantics of the
data is needed and must be supported by the overall infrastructure. Likewise,
information regarding typical access patterns for data in a given application
should be placed in the SSIO layers. Scientific applications often read and
analyze data in a structured fashion and failure to respect this pattern, or to
exploit it through prefetching, can significantly increase the overall
execution time of the application. Accurate prefetching can only be achieved
with access pattern analysis (\S\ref{sec:managing-data-life}). This requires an SSIO layer which is capable of
prefetching data in an asynchronous fashion to take the bytes on the file.

@@

\item Precision based re-organization. This is where the most significant bytes
	of the data are all grouped together from each object. This data
	generally will have a higher utility than the data with the least
	significant bytes. 
        %This process has the following three steps which
	%need to be efficiently implemented when data gets placed to or retrieved from the storage.
	The data needs to be re-arranged which involves potentially intensive 
        memory operations and needs to be done as much in situ as possible without
        involving the slow storage systems, and inter-processor communications.
	%At most this requires
	%two copies of an individual dataset in memory, and we will investigate
	%techniques to allocate and deallocate this memory if the user will
	%specify that the data will be overwritten after it is written to the
	%storage system. This is often the case for many of the  quantities
	%written from the simulation, but there are many cases where, for
	%example, we want to write all of the particles from a Particle In Cell
	%(PIC) simulation.  Since the particles will be used later in the
	%calculation, we need to duplicate the storage. Our observation with
	%working with the XGC1, GTC, and PiconGPU simulations 
	% is that we can temporary increase the
	%storage of the particles, and then release them since the temporary
	%arrays used in the calculations are often freed when a PIC iteration is
	%finished. 
%
\item Frequency-based re-organization.  Another common approach to classifying
	the importance of data is to re-organize data according to it frequency.
	This is commonly used in streaming data, and data reduction techniques
        such as those used JPEG-2000~\cite{jpeg2000}, which allow mechanisms
	to support spatial random access or region of interest
	access at varying degrees of granularity. It demonstrates the possibility of storing
	the same data using different quality. In relation to this project, this
        capability allows us to place the lowest frequency chunks in the fastest storage, and the
	highest frequency chunks on either the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out these pieces. 
%	This binning of data in frequency space can take as an input
%	(storage  {\color{red}ending )? CHECK-tsr} bandwidth knowledge, storage size limitations, and user
%	intentions of how long the majority of the data will be frequently
%	accessed. 
	 In order to fully take advantage of frequency based
	re-organization, the team has devised a scheme where the data pre-conditioner
	first sorts the data in bins with fixed length, and then wavelets and spline fits are used to
	re-organize the values.  This re-organization effectively produces
	a smoother data, allowing for much better reduction later.
\end{itemize}

### 2015-07-12 narrative/naming-service.tex

\subsection{Naming and Discovery Service}
\label{sec:naming-discovery}
\paragraph{Background:}
Given the well documented difficulties with POSIX-compliant naming and our
approach that decomposes what would be a single file into potentially large
number of objects, we must take a new approach to naming.  Data migration, as a
result of capacity pressures or user/system policies, further adds to the
complexity of managing naming and discovery. Our naming service must maintain
sufficient metadata information, track or be able to find objects as they move
across storage tiers, and facilitates data discovery. Since we need to also
support access through POSIX APIs for existing applications and system
integration, we must address how to bridge the drastic differences between the
decomposed objects of potentially varying fidelity
with a POSIX API-compatible interface.

Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm} offer a
strict caching approach for managing different storage capacities trading off
performance for capacity.  By maintaining a single namespace across all tiers,
it is possible to list a single directory view with files stored at different
tiers. While this approach to managing multi-tier storage works for whole files
that fit on single logical devices, it is far from ideal for scientific
simulations.

The overriding theme for this proposal of providing a cooperative relationship
between the user and the storage system extends into this area as well.
Negotiating the data quality/retrieval time trade-off requires additional
metadata.  Providing this functionality requires the naming service be able to
both return a list of data chunks and an indication of data quality. This
will be combined with a retrieval time estimate by the
middleware/application to determine what version should be retrieved.  This estimate
will be discussed elsewhere in this proposal.

In addition to the basic naming and data tracking operations, we will also need
to incorporate authorization capabilities. Sirocco currently integrates with a
Kerberos service for authentication. Given a capability ticket, a user can
access different objects as needed. This ticket structure offers protection
services typically offered on POSIX directories and files, but can do it at the
fork level instead. This allows a reduced quality verion to be available
for the general users while the high-quality version would be limited to the
data creator. This and other considerations for security must be incorporated
into the entire naming and metadata service.

While strongly encouraging data to stay in a particular type of storage tier,
ultimately it must be possible to flush this data to make room for other uses.
Sirocco offers a ``flush'' operation that forcibly migrates data to the
long-term resilience requirements, freeing space in the scarce resources. In
many, if not most cases, this will move data towards or actually on to a device
like tape, intended for long-term storage.

\paragraph{Approach:}
Our goal with this proposal is to offer a similar capability but with a finer
granularity leveraging the inner structure and utility of scientific data. It must
handle data discovery efficiently for cases where data is spread across several
storage tiers or some of those pieces may have migrated based on capacity
pressures or user/system policies. Instead of a single file, such as might
be used to store an entire timestep output for a simulation, we will
demonstrate a metadata service with naming at the application variable (e.g., particle) level, but
with an associated list of all of the variable pieces and potentially different
versions stored throughout the storage system. By shifting to a finer-grained
approach, we will enable more effective use of fast yet small storage tiers
and prioritize objects with high importance.  Traditional HSM stores an entire
file on a tier making room if insufficient space is available.  With this shift
to a partial variable granularity, a 1 PB output with 500 GB of ``high
interest'' data can limit this costly tier usage to just 500 GB, greatly
improving storage availability in the fast tiers and focus
performance/capacity costs to get the most science from the platform.  By
only placing ``high interest'' data in the highest tier, we will
effectively reduce the pressure forcing data migration, which hurts both writing
performance and the later data analytics required for generating scientific
insights.  Sirocco identifies a data segment using a
container/object/fork/address tuple that can have associated attributes. We will
add a special attribute indicating that this record is of ``high
interest'' as well adapting a migration mechanisms to take advantage of this attribute, if
present, to strongly encourage keeping the record in the highest storage tier as possible.

Meanwhile, we need to be able to locate or ``discover'' the data should it
move.  The challenge for discovery is that potentially, data will migrate from
where it is initially stored to a new location within the storage system.
Sirocco offers an ability to search for data that has moved as well as forcing
a particular resilience-based replica be the ``authoritative'' version
migrating the data to a particular location. We will investigate if the current
Sirocco functionality is capable of supporting the new operating modes we wish
to offer. Initial expectations suggest having bounded time guarantees for
finding data are critical for offering the QoS guarantees.
Some potential approaches can build on CRUSH~\cite{weil:ceph}
from Ceph to offer a map of where to search for data sequentially. Given our
multiple storage tiers, we would extend this using some mechanism to shift to
the next tier to continue searching because the requested data never made it to
this location. This new work will be an expansion of Sirocco's currently
planned features.

We plan to address scalability problems by continuing with the LWFS and Sirocco
model. LWFS and Sirocco have taken an approach similar to the pure object
stores, but with a focus on the HPC setting. They have abandoned a fully POSIX
compliant metadata service as the default model in favor of a
container/object/fork/address tuple for identifying data similar to those used
for pure object stores. By having a service that addresses the object
collections that comprise a single entity, such as a variable or timestep
output, sufficient metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support
traditional POSIX API calls even though the underlying storage system uses a
different model. Because we are not requiring the entire storage space be
addressable from a single tree root, we can offer multiple independent roots
using the Sirocco object storage. We will investigate how to make the naming
and discovery service scalable under the circumstances outlined above.

\paragraph{Related Work:}
Structurally, the traditional POSIX naming service offering a hierarchical
space consisting of directories and files may be maintained for backwards
compatibility. However, this view will not offer the same strict semantics
POSIX defines. We must break these strict semantics to address scalability
problems forced by the serialized access to a single source for creating and
accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have
worked to reduce this contention by reducing the serialized
scope to a single directory or subtree, or allowing a single process in a
collective file operation talk with the metadata service and distribute a
handle to other participants. While these approaches help, they do not address
this key scalability limitation of serialized access - even if the metadata
service is spread across multiple nodes. Instead, we will offer a short
duration to consistency POSIX view to offer the performance and consistency
requirements an exascale application demands.

@@

\paragraph{Challenge:}
The main challenge of incorporating the additional, rich metadata will be
joined by the additional challenge of coordinating with the other storage and
application layer services to offer the best access times possible for data
stored in the system. The developed metadata services that drive data
compression and subsetting operations must have easy, consistent, and ideally
non-blocking or locking access to this metadata service. We must investigate
how to build such a metadata and naming service that also incorporates and
maintains the additional metadata required to support our advanced
functionality.

>>>>

\subsection{Naming and Discovery Service}
\label{sec:naming-discovery}
\paragraph{Background:}
POSIX-compliant naming suffers from well documented scalability limitations due
to required operation serialization. Our approach would exascerbate this
problem because we decompose what would be a single file into potentially large
numbers of objects each broken into potentially multiple chunks to represent
data utility at a finer grained level to facilitate efficient storage usage.
Data migration, as a result of capacity pressures or user/system policies,
further adds to the naming management and discovery complexity.  Temporarily to
address performance requirements, we may make additional chunk replicas. In
some cases, such as with tape, having a highly compressed replica available in
a faster access medium near the tape system will offer efficient data probing
before forcing data retrieval from tape. In general, the performance replicas
will only exist as long as necessary to address performance requirements while
the tape overview version would serve as a metadata-like interface to the tape
archive. Since we must also support access through POSIX APIs for existing
applications and system integration, we must address how to bridge the drastic
differences between the objects and chunks of potentially varying fidelity with
a POSIX API-compatible interface.  Our naming service must maintain sufficient
metadata information to track or be able to find objects and chunks as they
move across storage tiers facilitating data discovery and be able to offer a
POSIX-compatible view for existing infrastructure integration.

The overriding theme for this proposal of providing a cooperative relationship
between the user and the storage system extends into this area as well.
Users should be able to understand the available data quality and approximate
retrieval times to enable making an appropriate trade-off for the current task.
Supporting this interaction requires additional metadata.  The naming service
must be able to both return a list of data chunks with their current
location and an indication of data quality. This will be combined with a
retrieval time estimate by the middleware/application to determine what version
should be retrieved.  This retrieval time estimate will be discussed elsewhere
in this proposal.

While, Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm}
seems similar to what we propose, that view only applies at the surface level.
HSM offers a strict caching approach for managing different storage capacities
trading off performance for capacity.  By maintaining a single namespace across
all tiers, it is possible to list a single directory view with files stored at
different tiers. HSM uses a whole file granularity. While this approach to
managing multi-tier storage works for whole files that fit on single logical
devices, it is far from ideal for scientific simulations. We will further
optimize the HSM approach to make the most efficient use of scarce storage
resources by only pulling the important/requested portions of a data set close
to the user. We propose to address this HSM limitation and support the required
complex naming and discovery metadata.

In addition to the basic naming and data tracking operations, we will also need
to incorporate authorization capabilities. Sirocco currently integrates with a
Kerberos service for authentication. Given a capability ticket, a user can
access different objects as needed. This ticket structure offers protection
services typically offered on POSIX directories and files, but can do it at the
fork level instead. This allows a reduced quality verion to be available
for the general users while the high-quality version would be limited to the
data creator. This and other considerations for security must be incorporated
into the entire naming and metadata service, particularly for
POSIX-compatibility.

While strongly encouraging data to stay in a particular storage tier,
ultimately it must be possible to flush this data to make room for other uses.
Sirocco offers a ``flush'' operation that forcibly migrates data to the
long-term resilience requirements, freeing space in the scarce resources. In
many, if not most cases, this will move data towards or actually on to a device
like tape, intended for long-term storage.

\paragraph{Approach:}
Our goal with this proposal is to offer an HSM-style capability but with a
finer granularity leveraging the inner structure and utility of scientific
data. It must handle data discovery efficiently for cases where data is spread
across several storage tiers or some of those pieces may have migrated based on
capacity pressures or user/system policies. Instead of a single file, such as
might be used to store an entire timestep output for a simulation, we will
demonstrate a metadata service with naming at the application variable (e.g.,
particle) level, but with an associated list of all of the variable pieces and
potentially different versions stored throughout the storage system (i.e., the
chunk list). By shifting to a finer-grained approach, we will enable more
effective use of fast yet small storage tiers and prioritize objects with high
importance.  With this shift to a partial variable granularity, a 1 PB output
with 500 GB of ``high interest'' data can limit this costly tier usage to just
500 GB, greatly improving storage availability in the fast tiers and focus
performance/capacity costs to get the most science from the platform.  By only
placing ``high interest'' data in the highest tier, we will effectively reduce
the pressure forcing data migration which hurts both writing performance and
the later data analytics required for generating scientific insights.  Sirocco
identifies a data segment using a container/object/fork/address tuple that can
have associated attributes. We will add a special attribute indicating that
this record is of ``high interest'' as well adapting a migration mechanisms to
take advantage of this attribute, if present, to strongly encourage keeping the
record in the highest storage tier as possible.

We also must be able to locate or ``discover'' the data should it move.  The
challenge for discovery is that potentially, data will migrate from where it is
initially stored to a new location within the storage system.  Sirocco offers
an ability to search for data that has moved as well as forcing a particular
resilience or quality-based replica be the ``authoritative'' version migrating
the data to a particular location. We will investigate if the current Sirocco
functionality is capable of supporting the new operating modes we wish to
offer. Initial expectations suggest having bounded time guarantees for finding
data are critical for offering the QoS guarantees or even strong best effort
support.  Some potential approaches can build on CRUSH~\cite{weil:ceph} from
Ceph to offer a map of where to search for data sequentially. Given our
multiple storage tiers, we would extend this using some mechanism to shift to
the next tier to continue searching because the requested data never made it to
this location. This new work will be an expansion of Sirocco's currently
planned features.

We plan to address scalability problems by continuing with the LWFS and Sirocco
model. LWFS and Sirocco have taken an approach similar to the pure object
stores, but with a focus on the HPC setting. They have abandoned a fully POSIX
compliant metadata service as the default model in favor of a
container/object/fork/address tuple for identifying data similar to those used
for pure object stores. By having a service that addresses the object
collections that comprise a single entity, such as a variable or timestep
output, sufficient metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support
traditional POSIX API calls even though the underlying storage system uses a
different model. Because we are not requiring the entire storage space be
addressable from a single tree root, we can offer multiple independent roots
using the Sirocco object storage. We will investigate how to make the naming
and discovery service scalable under the circumstances outlined above,
particularly to offer a coherent view given the potentially varying data
quality across different chunks.

\paragraph{Related Work:}
Structurally, the traditional POSIX naming service offering a hierarchical
space consisting of directories and files may be maintained for backwards
compatibility. However, this view will not offer the same strict semantics
POSIX defines. We must break these strict semantics to address scalability
problems forced by the serialized access to a single source for creating and
accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have
worked to reduce this contention by reducing the serialized
scope to a single directory or subtree or by allowing a single process in a
collective file operation talk with the metadata service and distribute a
handle to other participants. While these approaches help, they do not address
this key scalability limitation of serialized access - even if the metadata
service is spread across multiple nodes. Instead, we will offer a short
duration to consistency POSIX view to offer the performance and consistency
requirements an exascale application demands.

@@

\paragraph{Challenge:}
The main challenge of incorporating the additional, rich metadata will be
joined by the additional challenge of coordinating with the other storage and
application layer services to offer the best access times possible for data
stored in the system at available data quality levels. The developed metadata
services that drive data compression and subsetting operations must have easy,
consistent, and ideally non-blocking or locking access to this metadata
service. We must investigate how to build such a metadata and naming service
that also incorporates and maintains the additional metadata required to
support our advanced functionality.
